Privacy-Preserving Localisation Using Private Linear-Combination Aggregation\thanksref{footnoteinfo}

### 2020-11-04 main_automatica.tex
Initial switch to IEEE template (for Transactions on dependable and secure computing)

To show how $\rvec{i}'_{k,i}$ and $\mat{I}'_{k,i}$ can be formulated in a similar manner using \eqref{eqn:matrix_mult_weights}, we require the existance of a modified measurement model of the form
\begin{equation}
    \rv{z}'_{k,i} = h'_i(\rvec{x}_k)+\rv{v}'_{k,i}\,, \label{eqn:modified_measurement_model}
\end{equation}
where $\rv{z}'_{k,i}$ is the modified measurement, and noise $\rv{v}'_{k,i}$ is white and Gaussian. The approximation of $\rv{z}'_{k,i}$ and $\rv{r}_{k,i}$ from original measurements and variances $\rv{z}_{k,i}$ and $r_i$ will be shown in Section \ref{subsec:range_measurement_mod}.

@@

\subsection{Range Measurement Modification} \label{subsec:range_measurement_mod}
In Section \ref{subsec:observation_model}, we assumed the existance of modified sensor measurements $\rv{z}'_{k,i}$ with white Gaussian noise and variances $\rv{r}_{k,i}$ in the modified measurement model \eqref{eqn:modified_measurement_model}. In practice, conversion of range measurements $\rv{z}_{k,i}$ to $\rv{z}'_{k,i}$ is complicated by the noise term $\rv{v}_{k,i} \sim \mathcal{N}(0, r_i)$ in \eqref{eqn:measurement_model}. Squaring the range measurement produces
\begin{equation}
    \begin{split}
        \rv{z}_{k,i}^2 &= (h_i(\rvec{x}_k) + \rv{v}_{k,i})^2 \\
        &= h_i(\rvec{x}_k)^2 + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 \\
        &= h'_i(\rvec{x}_k) + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2\,,
    \end{split}
\end{equation}
with new noise term, $2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2$, now dependant on the measurement function $h_i$ and no longer white or Gaussian. We can compute the mean of the new noise term (a function of white Gaussian term $\rv{v}_{k,i}$) as
\begin{equation}
    \E{2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2} = r_i\, \label{eqn:modified_noise_mean}
\end{equation}
and define modified measurements as
\begin{equation} \label{eqn:modified_measurement}
    \begin{split}
        \rv{z}'_{k,i} &= \rv{z}_{k,i}^2 - r_i \\
        &= h_i(\rvec{x}_k)^2 + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 - r_i \\
        &= h'_i(\rvec{x}_k) + \rv{v}'_{k,i}\,,
    \end{split}
\end{equation}
with now zero-mean noise $\rv{v}'_{k,i} = 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 - r_i$. The noise in this case (again a function of white Gaussian term $\rv{v}_{k,i}$) has variance 
\begin{equation}
    \Var{\rv{v}'_{k,i}} = 4h_i(\rvec{x}_k)^2r_i + 2r_i^2
\end{equation}
and is dependant on $h_i(\rvec{x}_k)$. We also wish to approximate this noise using values available to the sensor, $\rv{z}_{k,i}$ and $r_i$, while providing a conservative estimate of the true variance so as not to affect the convergence of the EIF. While replacing $h_i(\rvec{x}_k)$ with $\rv{z}_{k,i}$ only provides a conservative estimate in the case
\begin{equation}
    \begin{split}
        &\rv{z}_{k,i} > h_i(\rvec{x}_k) \\
        \implies &\rv{z}_{k,i} - h_i(\rvec{x}_k) > 0 \\
        \implies &\rv{v}_{k,i} > 0\,,
    \end{split}
\end{equation}
which cannot be guaranteed, we can instead provide a conservative estimate with $95\%$ confidence, by shifting measurement $\rv{z}_{k,i}$ by two standard deviations $\sqrt{r_i}$. Thus, the modified measurement variance for sensor $i$ at time $k$ is conservatively approximated by
\begin{equation}
    \rv{r}_{k, i} = 4(\rv{z}_{k,i} + 2\sqrt{r_i})^2r_i + 2r_i^2\,. \label{eqn:modified_meaasurement_variance_estimate}
\end{equation}

In addition, we note that the distribution of noise in \eqref{eqn:modified_measurement} approaches a Gaussian when true ranges are sufficiently larger than measurement variances, as shown by
\begin{equation}
    \begin{split}
        &r_i << h_i(\rvec{x}_k) \\
        \implies &2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 - r_i \approx 2h_i(\rvec{x}_k)\rv{v}_{k,i} \\
        \implies &\rv{v}'_{k,i} \approx 2h_i(\rvec{x}_k)\rv{v}_{k,i}\,,
    \end{split} \label{eqn:modified_noise_approaches_gaussian}
\end{equation}
and conclude with the following assumption.
\begin{assum} \label{assum:measurement_size}
    Given property \eqref{eqn:modified_noise_approaches_gaussian}, and our $95\%$ confindence in the conservativness of variance estimates \eqref{eqn:modified_meaasurement_variance_estimate}, we assume that range measurements $\rv{z}_i$ are in general large enough to produce sufficiently Gaussian modified measurement noise $\rv{v}'_{k,i}$ and its variance estimated sufficiently conservative to be used in the EIF and measurement model \eqref{eqn:modified_measurement_model}.
\end{assum}

@@

    \State Measure $\rv{z}_{k,i}$
    \State Compute $\rv{z}_{k,i}'$ by \eqref{eqn:modified_measurement}
    \State Compute $\rv{r}_{k, i}$ by \eqref{eqn:modified_meaasurement_variance_estimate}

@@

From the leaked coefficient sums, we see that all information private to the sensors and broadcast to the navigator, namely their modified measurements $\rv{z}'_{k,i}$, variance estimates $\rv{r}_{k,i}$ and locations $\vec{s}_i$, are present only in the sums
\begin{equation}
    \sum_{i=1}^n\rv{z}'_{k,i}\,,\ \sum_{i=1}^n\rv{r}_{k,i}\,,\ \sum_{i=1}^ns_{x,i}\ \textrm{and}\ \sum_{i=1}^ns_{y,i}\,.\label{eqn:localisation_leakage}
\end{equation}
From these sums and the definitions \eqref{eqn:modified_measurement} and \eqref{eqn:modified_meaasurement_variance_estimate}, we conclude that the navigator can \textit{at most} learn the sums of sensor private data, which in practice corresponds to the average sensor modified measurements, estimated variances and locations.
\begin{thm}
    In the context of our localisation scheme, leakage of the incorporated LCAO secure encryption scheme corresponds to the leakage of average sensor private information, given by the sums in \eqref{eqn:localisation_leakage}.
\end{thm}

@@

Recalling Assumption \ref{assum:measurement_size}, modified measurement noise resembles a Gaussian distribution when measurements are sufficiently large. We considered $4$ layouts with varying sensor distances when running our simulations to observe differences in estimation error when measurements are of different sizes. The considered layouts have been plotted and named in Figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
\begin{center}
    \includegraphics{images/layouts.pdf}
\end{center}
\caption{The considered layouts.}
\label{fig:sim_layouts}
\end{figure}
To evaluate the accuracy of our method, we have run $100$ simulations, all with $50$ filter iterations, for each of the sensors layouts with sensor measurement variances set to $r_i=5$ simulation distance units. Due to the large encoding modulus $N$, no notable advantage is gained when choosing a low fractional precision factor $f$, thus, $f=2^{32}$ was chosen to ensure minimal loss of precision when compared to the floating-point computations in the unencrypted EIF. The plotted average RMSE of our proposed method as well as that of the unencrypted EIF filter can be seen in Figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
\begin{center}
    \includegraphics{images/layout_errors.pdf}
\end{center}
\caption{Filter error for different layouts.}
\label{fig:sim_layout_errors}
\end{figure}
From the results, it can be seen that sensor distance has little effect on the performance of the filter and that the measurement modification in Section \ref{subsec:range_measurement_mod} has almost no effect on the performance when compared to a normal EIF filter. The robustness of the privacy-preserving method is likely due to the rationality of Assumption \ref{assum:measurement_size}, the relative size of measurements to the sensor variances in our simulations, and the sufficiently large fractional precision.

@@

\section{Conclusion} \label{sec:conclusion}
To develop a model-based localisation algorithm which preserves navigator and sensor privacy, we first defined the novel notion of an LCAO secure encryption scheme and gave a provably secure implementation based on the Joye-Libert aggregation scheme. This scheme was then used in our proposed privacy-preserving EIF, where sensors compute measurement information homomorphically such that it can be privately used by a navigator. Our privacy-preserving estimation method may find uses in a variety of untrusted distributed localisation environments including airspaces and autonomous vehicle networks as well as those where alternative measurement models satisfying our defined linearity requirements can be applied. Possible future work in this topic includes the expanding of the LCAO security notion to \textit{ensure} that the same weights are broadcast to all sensors, the relaxing of the Gaussian modified noise assumption and the exploring of implications behind an active sensor attacker model.

>>>>

To show how $\rvec{i}'_{k,i}$ and $\mat{I}'_{k,i}$ can be formulated in a similar manner using \eqref{eqn:matrix_mult_weights}, we require the existance of a modified measurement model of the form
\begin{equation}
    \rv{z}'_{k,i} = h'_i(\rvec{x}_k)+\rv{v}'_{k,i}\,, \label{eqn:modified_measurement_model}
\end{equation}
where $\rv{z}'_{k,i}$ is the modified measurement, and noise $\rv{v}'_{k,i}$ is zero-mean with known variance. The approximation of $\rv{z}'_{k,i}$ and its noise variance $\rv{r}_{k,i}$ from original measurements and variances $\rv{z}_{k,i}$ and $r_i$ will be shown in Section \ref{subsec:range_measurement_mod}.

@@

\subsection{Range Measurement Modification} \label{subsec:range_measurement_mod}
In Section \ref{subsec:observation_model}, we assumed the existance of a measurement model \eqref{eqn:modified_measurement_model} where modified sensor measurements $\rv{z}'_{k,i}$ had zero-mean noises with variances $\rv{r}_{k,i}$. In practice, conversion of range measurements $\rv{z}_{k,i}$ to $\rv{z}'_{k,i}$ is complicated by the noise term $\rv{v}_{k,i} \sim \mathcal{N}(0, r_i)$ in \eqref{eqn:measurement_model}. Squaring the range measurement produces
\begin{equation}
    \begin{split}
        \rv{z}_{k,i}^2 &= (h_i(\rvec{x}_k) + \rv{v}_{k,i})^2 \\
        &= h_i(\rvec{x}_k)^2 + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 \\
        &= h'_i(\rvec{x}_k) + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2\,,
    \end{split}
\end{equation}
with new noise term, $2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2$, now dependant on the measurement function $h_i$ and no longer zero-mean or Gaussian. We can compute the mean of the new noise term (a function of the Gaussian term $\rv{v}_{k,i}$) as
\begin{equation}
    \E{2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2} = r_i\, \label{eqn:modified_noise_mean}
\end{equation}
and define modified measurements as
\begin{equation} \label{eqn:modified_measurement}
    \begin{split}
        \rv{z}'_{k,i} &= \rv{z}_{k,i}^2 - r_i \\
        &= h_i(\rvec{x}_k)^2 + 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 - r_i \\
        &= h'_i(\rvec{x}_k) + \rv{v}'_{k,i}\,,
    \end{split}
\end{equation}
with now zero-mean noise $\rv{v}'_{k,i} = 2h_i(\rvec{x}_k)\rv{v}_{k,i} + \rv{v}_{k,i}^2 - r_i$. The noise in this case (again a function of white Gaussian term $\rv{v}_{k,i}$) has variance 
\begin{equation}
    \Var{\rv{v}'_{k,i}} = 4h_i(\rvec{x}_k)^2r_i + 2r_i^2
\end{equation}
and is dependant on $h_i(\rvec{x}_k)$. To use the modified measurement function \eqref{eqn:modified_measurement} with an EIF, we require an estimate of its noise variance using only the values available to the sensor, $\rv{z}_{k,i}$ and $r_i$. A conservative estimate of the variance (\textit{i.e.} a larger estimate resulting in less confidence in measurements) is desirable to avoid increasing the likeliness of filter divergence. While replacing $h_i(\rvec{x}_k)$ with $\rv{z}_{k,i}$ only provides a conservative estimate when
\begin{equation}
    \begin{split}
        &\rv{z}_{k,i} > h_i(\rvec{x}_k) \\
        \implies &\rv{z}_{k,i} - h_i(\rvec{x}_k) > 0 \\
        \implies &\rv{v}_{k,i} > 0\,,
    \end{split}
\end{equation}
and cannot be guaranteed, we can instead provide a conservative estimate with $95\%$ confidence, by shifting measurement $\rv{z}_{k,i}$ by two standard deviations $\sqrt{r_i}$. Thus, the modified measurement variance for sensor $i$ at time $k$ is conservatively approximated by
\begin{equation}
    \rv{r}_{k, i} = 4(\rv{z}_{k,i} + 2\sqrt{r_i})^2r_i + 2r_i^2\,. \label{eqn:modified_measurement_variance_estimate}
\end{equation}
Together, the model and variance, \eqref{eqn:modified_measurement} and \eqref{eqn:modified_measurement_variance_estimate}, provide a suitable substitute for the required measurement model \eqref{eqn:modified_measurement_model} and can be used to achieve the desired linear properties using only the available range sensors.

@@

    \State Measure $\rv{z}_{k,i}$
    \State Compute $\rv{z}_{k,i}'$ by \eqref{eqn:modified_measurement}
    \State Compute $\rv{r}_{k, i}$ by \eqref{eqn:modified_measurement_variance_estimate}

@@

From the leaked coefficient sums, we see that all information private to the sensors and broadcast to the navigator, namely their modified measurements $\rv{z}'_{k,i}$, variance estimates $\rv{r}_{k,i}$ and locations $\vec{s}_i$, are present only in the sums
\begin{equation}
    \sum_{i=1}^n\rv{z}'_{k,i}\,,\ \sum_{i=1}^n\rv{r}_{k,i}\,,\ \sum_{i=1}^ns_{x,i}\ \textrm{and}\ \sum_{i=1}^ns_{y,i}\,.\label{eqn:localisation_leakage}
\end{equation}
From these sums and the definitions \eqref{eqn:modified_measurement} and \eqref{eqn:modified_measurement_variance_estimate}, we conclude that the navigator can \textit{at most} learn the sums of sensor private data, which in practice corresponds to the average sensor modified measurements, estimated variances and locations.
\begin{thm}
    In the context of our localisation scheme, leakage of the incorporated LCAO secure encryption scheme corresponds to the leakage of average sensor private information, given by the sums in \eqref{eqn:localisation_leakage}.
\end{thm}

@@

Recalling the dependence of the modified measurement noise on the true measurement in \eqref{eqn:modified_measurement}, the noise distribution of the model changes with varying sizes of measurements. We considered four layouts with varying sensor distances when running our simulations to observe differences in estimation error when measurements are of different sizes. The considered layouts have been plotted and named in Figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
\begin{center}
    \includegraphics{images/layouts.pdf}
\end{center}
\caption{The considered layouts.}
\label{fig:sim_layouts}
\end{figure}
To evaluate the accuracy of our method, we have run $100$ simulations, all with $50$ filter iterations, for each of the sensors layouts with sensor measurement variances set to $r_i=5$ simulation distance units. Due to the large encoding modulus $N$, no notable advantage is gained when choosing a low fractional precision factor $f$, thus, $f=2^{32}$ was chosen to ensure minimal loss of precision when compared to the floating-point computations in the unencrypted EIF. The plotted average RMSE of our proposed method as well as that of the unencrypted EIF filter can be seen in Figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
\begin{center}
    \includegraphics{images/layout_errors.pdf}
\end{center}
\caption{Filter error for different layouts.}
\label{fig:sim_layout_errors}
\end{figure}
From the results, it can be seen that sensor distance has little effect on the performance of the filter and that the measurement modification from Section \ref{subsec:range_measurement_mod} has a negligible effect when compared to a normal EIF filter. The robustness of the privacy-preserving method can be attributed to the conservativeness of variance estimates in \eqref{eqn:modified_measurement_variance_estimate} and the sufficiently large fractional precision parameter.

@@

\section{Conclusion} \label{sec:conclusion}
To develop a model-based localisation algorithm which preserves navigator and sensor privacy, we first defined the novel notion of an LCAO secure encryption scheme and gave a provably secure implementation based on the Joye-Libert aggregation scheme. This scheme was then used in our proposed privacy-preserving EIF, where sensors compute measurement information homomorphically such that it can be privately used by a navigator. Our privacy-preserving estimation method may find uses in a variety of untrusted distributed localisation environments including airspaces and autonomous vehicle networks as well as those where alternative measurement models satisfying our defined linearity requirements can be applied. Possible future work in this topic includes the expanding of the LCAO security notion to \textit{ensure} that the same weights are broadcast to all sensors, the application of LCAO secure schemes to alternative measurement models and the exploring of implications behind an active sensor attacker model.
